---
title: "Data Wrangling Exercise 2"
author: "Michael Moran"
date: "July 11, 2017"
output: html_document
---

#Data Wrangling Exercise 2

```{r}
library(dplyr)
library(readr)
library(tidyr)

titanic3 <- read_csv("~/Foundations of Data Science/titanic_original.csv")
tbl_df(titanic3)
```


#Step One: Missing values for embarkment

```{r}
titanic_embarked <- titanic3 %>% mutate(embarked = case_when(.$embarked == c('is.na') ~ 'S', TRUE ~ .$embarked))
```


#Step Two: Missing values for age

```{r}
mean_age <- mean(titanic_embarked$age, na.rm = TRUE)
titanic_age <- titanic_embarked %>% replace_na(list(age = mean_age))
```


#Using the mean for age may be more affected by outliers in the dataset, so using the median would be an alternative to the mean without that problem. 

#Step Three: Fill the lifeboat column

```{r}
titanic_lifeboat <- titanic_age %>% replace_na(list(boat = 'None'))
```


#Step Four: 

```{r}
titanic_cabin <- titanic_lifeboat %>% mutate(has_cabin_number = if_else(is.na(cabin), 0, 1))
```


#A missing value here likely indicated that the individual did not survive the disaster.

#Step Five: Write cleaned data to csv

```{r}
write.csv(titanic_cabin, "titanic_clean.csv")
```

